### 2023-06-21 bachelorthesis/23-iontempgrad.tex
#39: Correction Anna-Maria

The ITG-driven instability is present when a sufficiently large ion temperature gradient occurs and is driven by the so-called bad curvature of the outboard low field side of the tokamak. The mechanism is briefly explained with the help of Fig. \ref{fig:ITG} and Ref. \citenum{Beer_PHD, Casson_PHD, Dannert_PHD}. An ion temperature perturbation ($T^\pm \gtrless 0$ \textcolor{blue}{blue} and \textcolor{red}{red} cells) on the outboard low field side of the tokamak causes the modulation of curvature drift and $\nabla B$ drift with velocity $\vect{v}_\mathrm{d}$. The divergence of the velocity $\vect{v}_\mathrm{d}$ is therefore not zero which results in compression of the plasma ($n_\pm \gtrless 0$, \textcolor{gray}{gray} cells). This relates to the perturbation of the potential ($\phi^\pm  \gtrless 0$) in the adiabatic electron response. As stated in the equation for the electric field ($\vect{E} = -\nabla \phi$) the perturbation of the potential $\phi$ causes an electric field $\vect{E}$ which results in a $\exb$ drift with velocity $\vect{v}_E$. The $\exb$ drift direction transports hot plasma in the outer hot region of perturbation ($T^+ > 0$, \textcolor{red}{red} cell) and cooler plasma into inner cold region perturbation ($T^- < 0$, \textcolor{blue}{blue} cell) which reinforce the process and instability grows.
\newpage
The described mechanism causes perturbation on the outboard low field side of the tokamak this region has a \enquote{bad curvature} and instability happens there. On the inboard high field side the opposing $\nabla B$ and $\nabla T$ results in a decay of perturbation which is referred to as \enquote{good curvature}. Furthermore, such turbulent structures are called \textit{eddies} which have a typical radial scale of several Larmor radii. \cite{Newins2006}
The initial perturbation is growing if the normalized temperature gradient overcomes a critical value $R/L_\mathrm{T,c}$. For nonlinear turbulence with Cyclone Base Case parameters and circular concentric flux-surfaces the critical value is given by $R/L_\mathrm{T,c} \sim 6$. \cite{Dimits2000, Isliker2010}

>>>>

The ITG-driven instability is present when a sufficiently large ion temperature gradient occurs and is driven by the so-called bad curvature of the outboard low field side of the tokamak. The mechanism is briefly explained with the help of Fig. \ref{fig:ITG} and Ref. \citenum{Beer_PHD, Casson_PHD, Dannert_PHD}. An ion temperature perturbation ($T^\pm \gtrless 0$ \textcolor{blue}{blue} and \textcolor{red}{red} cells) on the outboard low field side of the tokamak causes the modulation of curvature drift and $\nabla B$ drift with velocity $\vect{v}_\mathrm{d}$. The divergence of the velocity $\vect{v}_\mathrm{d}$ is therefore not zero which results in compression of the plasma ($n_\pm \gtrless 0$, \textcolor{gray}{gray} cells). This relates to the perturbation of the potential ($\phi^\pm  \gtrless 0$) in the adiabatic electron response. As stated in the equation for the electric field ($\vect{E} = -\nabla \phi$) the perturbation of the potential $\phi$ causes an electric field $\vect{E}$ which results in a $\exb$ drift with velocity $\vect{v}_E$. The $\exb$ drift direction transports hot plasma in the outer hot region of perturbation ($T^+ > 0$, \textcolor{red}{red} cell) and cooler plasma into inner cold region perturbation ($T^- < 0$, \textcolor{blue}{blue} cell) which reinforce the process and instability grows.
\newpage
The described mechanism causes perturbation on the outboard low field side of the tokamak this region has a \enquote{bad curvature} and instability happens there. On the inboard high field side the opposing $\nabla B$ and $\nabla T$ results in a decay of perturbation which is referred to as \enquote{good curvature}. Furthermore, such turbulent structures are called \textit{eddies} which have a typical radial scale of several Larmor radii. \cite{Newins2006}
The initial perturbation is growing if the normalized temperature gradient overcomes a critical value $R/L_{T,\mathrm{c}}$. For nonlinear turbulence with Cyclone Base Case parameters and circular concentric flux-surfaces the critical value is given by $R/L_{T,\mathrm{c}} \sim 6$. \cite{Dimits2000, Isliker2010}

### 2023-06-21 bachelorthesis/24-zonalflows.tex
#39: Correction Anna-Maria

Furthermore, zonal flows are connected with the radially alternating zonal potential, which introduces a velocity shear, i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence \cite{Biglari1990, Dimits2000}, especially the ion temperature gradient driven turbulence \cite{Nakata2012, Makwana2014, Maeyama2014, Whelan2018, Whelan2019}. The nonlinear generation mechanism causes a shift, also known as Dimits shift \cite{Dimits2000}, of the critical temperature gradient $R/L_\mathrm{T,c}$ which leads to a development of ITG-driven turbulence at higher $\rlt$ compared to the linear threshold. The shearing process tilts and stretches turbulent eddies, caused by the shear flow, until they break into smaller eddies with reduced radial correlation length [Fig. \ref{fig:zonalflow}],  characterizing the shearing as a decorrelation process. Turbulence will then be reduced because of this process. \cite{Biglari1990,Diamond2005,Burnell1997} \bigskip

>>>>

Furthermore, zonal flows are connected with the radially alternating zonal potential, which introduces a velocity shear, i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence \cite{Biglari1990, Dimits2000}, especially the ion temperature gradient driven turbulence \cite{Nakata2012, Makwana2014, Maeyama2014, Whelan2018, Whelan2019}. The nonlinear generation mechanism causes a shift, also known as Dimits shift \cite{Dimits2000}, of the critical temperature gradient $R/L_{T,\mathrm{c}}$ which leads to a development of ITG-driven turbulence at higher $\rlt$ compared to the linear threshold. The shearing process tilts and stretches turbulent eddies, caused by the shear flow, until they break into smaller eddies with reduced radial correlation length [Fig. \ref{fig:zonalflow}],  characterizing the shearing as a decorrelation process. Turbulence will then be reduced because of this process. \cite{Biglari1990,Diamond2005,Burnell1997} \bigskip

### 2023-06-21 bachelorthesis/43-heatfluxthreshold.tex
#39: Bachelor thesis finished
#39: Correction Anna-Maria

In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{Peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with
\begin{gather*}
	\rlt \in [6.0,~6.2,~6.4]~.
\end{gather*}
In the case of $\rlt = 6.2$ turbulence suppression is observed for $t > 11000\,R/\vth$, while stationary turbulence during the entire simulation time trace of $12000\,R/\vth$ is found for $\rlt = 6.4$ [Fig. \ref{fig:eflux-6.0-6.2-6.4-comparison}].
The finite heat flux threshold, hence, is
\begin{gather*}
	\boxed{\rlt|_\mathrm{finite} = 6.3 \pm 0.1}
\end{gather*}
in accordance to Ref. \citenum{Peeters2016}.
Although the initial quasi-stationary turbulence in the former case is significantly longer compared to the $\rlt = 6.2$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = 3$ establishes. 
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern size of $\sim 57.20 - 76.27\,\rhoth$.

\includegraphicsHere{Comparison/Gradient-Length/S6_rlt6.0-6.2-6.4_boxsize3x3_Ns16_Nvpar48_Nmu9_eflux_comparison.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for different gradient lengths $\rlt$
}{fig:eflux-6.0-6.2-6.4-comparison}{}

>>>>

In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{Peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with
\begin{gather*}
	\rlt \in [6.0,~6.2,~6.4]~.
\end{gather*}
In the case of $\rlt = 6.2$ turbulence suppression is observed for $t > 11000\,R/\vth$, while stationary turbulence during the entire simulation time trace of $12000\,R/\vth$ is found for $\rlt = 6.4$ [Fig. \ref{fig:eflux-6.0-6.2-6.4-comparison}].
The finite heat flux threshold, hence, is
\begin{gather*}
	\boxed{\rlt|_\mathrm{finite} = 6.3 \pm 0.1}
\end{gather*}
in accordance with Ref. \citenum{Peeters2016}.
Although the initial quasi-stationary turbulence in the former case is significantly longer compared to the $\rlt = 6.2$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = 3$ establishes. 
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern size of $\sim 57 - 76\,\rhoth$.

\includegraphicsHere{Comparison/Gradient-Length/S6_rlt6.0-6.2-6.4_boxsize3x3_Ns16_Nvpar48_Nmu9_eflux_comparison.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for different\\gradient lengths $\rlt$
}{fig:eflux-6.0-6.2-6.4-comparison}{}
